---
title: "Single-case rehabilitation reasoning with PhysioRehab"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Single-case rehabilitation reasoning with PhysioRehab}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

PhysioRehab re-centres physiological measurement around the way rehabilitation
reasons about the *individual*: it anchors measures and goals to the WHO
International Classification of Functioning (ICF), detects change with single-case
methods, keeps the clinician in the loop, and auto-drafts an ICF-structured
progress note.

This vignette walks the whole path on a **synthetic** case. The numbers below are
illustrative and are **not clinical advice**: PhysioRehab is a research tool for
measurement, documentation and clinician-in-the-loop reasoning support, not a
medical device, and final judgement always rests with the clinician.

```{r setup}
library(PhysioRehab)
```

## 1. A synthetic single case

`simulate_stroke_gait_case()` builds one patient's episode: an AB single-case
design with three baseline sessions and nine intervention sessions, recording
comfortable gait speed alongside ICF qualifiers and a Goal Attainment Scaling
(GAS) rating.

```{r episode}
ep <- simulate_stroke_gait_case()
ep
```

An episode is a container of repeated measures; you can also build one by hand
with `new_rehab_episode()` and `add_session()`, attaching a personal goal with
`rehab_goal()`.

## 2. Detect change defensibly

Rehabilitation compares a person to their own baseline, so change is detected
with distribution-free single-case statistics plus a minimal clinically
important difference (MCID). `sced_analyze()` returns the non-overlap of all
pairs (NAP), Tau, the baseline-vs-latest change, and whether it clears the MCID.

```{r sced}
sced <- sced_analyze(ep, measure = "gait_speed", mcid = 0.16)
sced[c("baseline_mean", "latest", "delta", "nap", "exceeds_mcid")]
```

The ICF qualifiers (0 = no problem ... 4 = complete problem; lower is better) are
tracked per code. Improvement means the qualifier *falls*:

```{r icf}
icf_change(ep, "d450")   # walking
icf_change(ep, "b735")   # muscle tone / spasticity
```

## 3. Keep the clinician in the loop

The clinician states an expectation; `evaluate_hypothesis()` reports, item by
item, whether the evidence supports it -- and honestly flags what it does not.

```{r reasoning}
hyp <- rehab_hypothesis(
  "task-oriented training improves gait",
  expect_speed_mcid = TRUE,
  expect_icf = c(d450 = 1, b735 = 1)
)
reasoning <- evaluate_hypothesis(ep, hyp, sced)
reasoning$overall
for (it in reasoning$items) {
  mark <- if (is.na(it$supported)) "-" else if (it$supported) "OK" else "NO"
  cat(sprintf("[%s] %s\n", mark, it$claim))
}
```

## 4. Visualise the trajectory

```{r plot, fig.width = 7, fig.height = 4.2}
plot_trajectory(ep, mcid = 0.16, ylab = "comfortable gait speed (m/s)")
```

## 5. Draft the progress note

The adoption hook is workflow: the tool drafts the ICF-structured progress note
the therapist already has to write, so the measurement and analysis ride along.

```{r note}
note <- draft_progress_note(ep, sced, reasoning)
cat(substr(note, 1, 400))
```

`rehab_workflow()` runs all of the above in one call (single-case analysis,
hypothesis evaluation, trajectory plot and drafted note), writing the artefacts
into `out_dir`:

```{r workflow}
res <- rehab_workflow(ep, hyp, out_dir = tempdir())
names(res)
```

## 6. Capacity vs performance (cross-modal ICF constructs)

Rehabilitation lives in the gap between what a person *can* do in the clinic
(capacity) and what they actually *do* at home (performance). An `icf_construct`
unifies heterogeneous measures on one ICF code by mapping each to the 0--4
qualifier via ascending cut-points.

```{r construct}
cc <- new_icf_construct("d450")
cc <- add_icf_measure(cc, "barthel_walk", value = 10,
                      qualifier_type = "capacity", modality = "scale",
                      cutpoints = c(2, 5, 8, 12))
cc <- add_icf_measure(cc, "steps_per_day", value = 1800,
                      qualifier_type = "performance", modality = "sensor",
                      cutpoints = c(1000, 3000, 6000, 9000))
capacity_performance_gap(cc)
```

## 7. What the ecosystem can measure

`adl_coverage_summary()` reports, for each activity of daily living, how many
modalities can measure it and whether both the capacity and performance
qualifiers are covered.

```{r coverage}
head(adl_coverage_summary())
```

## Where to go next

- Single-case statistics come from **PhysioAppKit** (the engine behind
  `sced_analyze()`).
- The ICF ontology is served by **PhysioAnnotationHub** (`icf_for_measure()`).
- Real signals feed the episode through **PhysioMoCap**
  (`add_session_from_mocap()`) and **PhysioEMG** (`add_session_from_emg()`).
- `?PhysioRehab` lists every entry point grouped by task.
